Caring and Feeding the Wild Caught Data

### 2021-03-14 New_Wages.Rmd
Add getting and tidy data section.
Change the threshold to be 0.1

## getting the highest year of completed education

@@

@@

hours_wages <- left_join(rates_all, 
                         hours_all, 
                         by = c("id", "year", "job")) %>%

# set the 0 value in rate_per_hour as NA 
  mutate(rate_per_hour = ifelse(rate_per_hour == 0, NA,
                                rate_per_hour))
```

# calculate number of jobs that a person has in one year
no_job <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  group_by(id, year) %>%
  summarise(no_jobs = length(rate_per_hour))

# filter the observations with available rate per hour
eligible_wages <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  left_join(no_job, by = c("id", "year"))

@@

mean_hourly_wage <- 
  eligible_wages %>%
  group_by(id, year) %>%
  #calculate the weighted mean if the number of jobs > 1
  mutate(wages = ifelse(no_jobs == 1, rate_per_hour/100,
                        weighted.mean(rate_per_hour, hours_work, na.rm = TRUE)/100)) %>%
  #give the flag if it the weighted mean
  mutate(flag1 = ifelse(!is.na(wages) & no_jobs != 1, 1,
                        0)) %>%
  #calculate the arithmetic mean for the na
  mutate(wages = ifelse(is.na(wages), mean(rate_per_hour)/100,
                        wages))

>>>>

## getting the highest year of completed education ever

@@

```{r}
hours_wages <- left_join(rates_all, 
                         hours_all, 
                         by = c("id", "year", "job")) %>%

# set the 0 value in rate_per_hour as NA 
  mutate(rate_per_hour = ifelse(rate_per_hour == 0, NA,
                                rate_per_hour))

hours_wages <- hours_wages %>%
  mutate(rate_per_hour_adj = ifelse(as.numeric(year) < 1994 & rate_per_hour > 1000, 1000,
                                    ifelse(as.numeric(year) >= 1994 & rate_per_hour > 5000, 5000, rate_per_hour)))
```

@@

# calculate number of jobs that a person has in one year
no_job <- hours_wages %>%
  filter(!is.na(rate_per_hour_adj)) %>%
  group_by(id, year) %>%
  summarise(no_jobs = length(rate_per_hour_adj))

# filter the observations with available rate per hour
eligible_wages <- hours_wages %>%
  filter(!is.na(rate_per_hour_adj)) %>%
  left_join(no_job, by = c("id", "year"))

@@

mean_hourly_wage <- 
  eligible_wages %>%
  group_by(id, year) %>%
  #calculate the weighted mean if the number of jobs > 1
  mutate(wages = ifelse(no_jobs == 1, rate_per_hour_adj/100,
                        weighted.mean(rate_per_hour_adj, hours_work, na.rm = TRUE)/100)) %>%
  #give the flag if it the weighted mean
  mutate(flag1 = ifelse(!is.na(wages) & no_jobs != 1, 1,
                        0)) %>%
  #calculate the arithmetic mean for the na
  mutate(wages = ifelse(is.na(wages), mean(rate_per_hour_adj)/100,
                        wages))

### 2021-03-14 report_alt.Rmd
Change the threshold to be 0.1

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE,
                      cache = TRUE, 
                      cache.path = "cache/",
                      fig.path = "figures/")
library(tidyverse)
library(brolgar)
library(patchwork)
library(kableExtra)
library(MASS)
library(janitor)
library(DiagrammeR)
```

@@

Just like the living things, the usage of the term "wild" could also be applied in the data domain. Wild data is the form of data that could be acquired from various source and might be generated from experiments, surveys, or may be from sensors or social media. Wild data also often associated with "Open Data", means that the data are freely accessible, modified, and shared by anyone for any purpose [@opendata].

Despite its traits, wild data often have an issue with its quality in the sense that it is often not tidy and clean. Hence, it should be tidied and cleaned before further analysis to prevent inappropriate usage of statistical method that may lead to a misleading result [@HuebnerMariannePhD2016Asat]. @HuebnerMariannePhD2016Asat also argued that data tidying and cleaning is the first part of Initial Data Analysis (IDA). The second and third steps are to explore the data properties and to document and report the process for the later formal analysis. This second step is the same with Chatfield's perspective on the goal of IDA [@Chatfield1985TIEo], which is to get the "feel" of the data. In other word, IDA could be regarded as a way to assess the data quality.

@@

# Create a data set for the high school dropouts cohort
wages_hs_dropout <- wages_hs2020 %>%
  mutate(dob = 1979 - age_1979,
         age_hgc = yr_hgc - dob) %>%
  filter((hgc %in% c("9TH GRADE",
                     "10TH GRADE",
                     "11TH GRADE")) |
           (hgc == "12TH GRADE" &
              age_hgc >= 19)) %>%
  dplyr::select(-dob,
         -age_hgc)

@@

```{r nrow, echo FALSE}
a = nrow(categories_qnames)
b = nrow(full_demographics)
d = eligible_wages %>%
  group_by(id) %>%
  count(id) %>%
  nrow()
n_hgc = wages_hs2020 %>%
  group_by(id) %>%
  count(id) %>%
  nrow()
n_obs_hgc = nrow(wages_hs2020)
n_obs_hgc_pred = filter(wages_hs2020, is_pred == TRUE) %>%
  nrow()

@@

```{r flowchart, fig.cap="The stages of data filtering from the raw data to get three datasets that are contained in an R package called 'yowie', 'n' means the number of ID, while 'n_obs' means the number of observations."}

grViz("digraph flowchart {
      node [fontname = Helvetica, shape = rectangle]        
      tab1 [label = '@@1']
      tab2 [label = '@@2']
      tab3 [label = '@@3']
      tab4 [label = '@@4']
      tab5 [label = '@@5']
      tab6 [label = '@@6']
      tab7 [label = '@@7']

@@

      [1]: paste0('NLSY79, n = ', a)
      [2]: paste0('Extract demographic variable, n = ', b)
      [3]: paste0('Exclude ', a - d, ' IDs whose hourly rate is missing')
      [4]: paste0('Eligible ID, n = ', d)
      [5]: paste0('Cohort whose hgc is up to 12th grade and \\n participated at least 5 years in the survey, \\n n = ', n_hgc, ', n_obs = ', n_obs_hgc, ' \\n (', n_obs_hgc_pred, ' observations are predicted value)')
      [6]: paste0('High school dropouts cohort, \\n n = ', n_do, ', n_obs = ', n_obs_do, ' \\n (', n_obs_do_pred, ' observations are predicted value)')
      [7]: 'yowie Package'
      ")
```

@@

>>>>

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE,
                      cache = TRUE, 
                      cache.path = "cache/",
                      fig.path = "figures/")
library(tidyverse)
library(brolgar)
library(patchwork)
library(kableExtra)
library(MASS)
library(janitor)
library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)
library(knitr)
library(webshot)
```

@@

The idea of "Open Data" is that the data are freely accessible, modified, and shared by anyone for any purpose [@opendata]. Due too its openness, Open Data are sometimes used and modified to be incorporated as the data in statistical text book or in research. Despite its traits, open data often have an issue with its quality in the sense that it is often not tidy and clean. Hence, it should be tidied and cleaned before further analysis to prevent inappropriate usage of statistical method that may lead to a misleading result [@HuebnerMariannePhD2016Asat].

@HuebnerMariannePhD2016Asat also argued that data tidying and cleaning is the first part of Initial Data Analysis (IDA). The second and third steps are to explore the data properties and to document and report the process for the later formal analysis. This second step is the same with Chatfield's perspective on the goal of IDA [@Chatfield1985TIEo], which is to get the "feel" of the data. In other word, IDA could be regarded as a way to assess the data quality.

@@

# Create a data set for the high school dropouts cohort
wages_hs_dropout <- wages_hs2020 %>%
  mutate(dob = 1979 - age_1979,
         age_hgc = yr_hgc - dob) %>%
  filter((hgc %in% c("9TH GRADE",
                     "10TH GRADE",
                     "11TH GRADE")) |
          (hgc == "12TH GRADE" &
              age_hgc > 19)) %>%
  filter(age_1979 <= 17,
         gender == "MALE") %>%
  dplyr::select(-dob,
         -age_hgc)

@@

```{r nrow, echo=FALSE}
a = nrow(categories_qnames)
b = nrow(full_demographics)
d = eligible_wages %>%
  group_by(id) %>%
  count(id) %>%
  nrow()
n_hgc = wages_hs2020 %>%
  group_by(id) %>%
  count(id) %>%
  nrow()
n_obs_hgc = nrow(wages_hs2020)
n_obs_hgc_pred = filter(wages_hs2020, is_pred == TRUE) %>%
  nrow()

@@

```{r flowchart, echo=FALSE, fig.cap="The stages of data filtering from the raw data to get three datasets that are contained in an R package called yowie, n means the number of ID, while n_obs means the number of observations.", fig.height=3, fig.width=3}

fc <- grViz("digraph flowchart {
      node [fontname = Helvetica, shape = rectangle]        
      tab1 [label = '@@1']
      tab2 [label = '@@2']
      tab3 [label = '@@3']
      tab4 [label = '@@4']
      tab5 [label = '@@5']
      tab6 [label = '@@6']
      tab7 [label = '@@7']

@@

      [1]: paste0('NLSY79, n = ', a)
      [2]: paste0('Extract demographic variable, n = ', b)
      [3]: paste0('Exclude ', a - d, ' IDs whose hourly rate is missing')
      [4]: paste0('Eligible ID, n = ', d)
      [5]: paste0('Cohort whose hgc is up to 12th grade and \\n participated at least 5 years in the survey, \\n n = ', n_hgc, ', n_obs = ', n_obs_hgc, ' \\n (', n_obs_hgc_pred, ' observations are predicted value)')
      [6]: paste0('High school dropouts cohort, \\n n = ', n_do, ', n_obs = ', n_obs_do, ' \\n (', n_obs_do_pred, ' observations are predicted value)')
      [7]: 'yowie Package'
      ",
      height = 200)

fc
```

@@

- Data provider should provide validation rules for data user
